From Real-time Attention Assessment to ``With-me-ness'' in Human-Robot Interaction

### 2015-09-17 paper.tex
Updated experimental setting figure
Formally introduce with-me-ness

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.7\columnwidth]{drawSetup}
    \caption{\small \textbf{Experimental setting} Graphical representation of the set-up and the possible children focuses of attention in the scene, in red.}
    \label{drawSetup}
\end{figure}

@@

\begin{figure*}
    \centering
    \includegraphics[width=1.81\columnwidth]{realCaptured}
    \caption{\small Real VFoA of subject 4 accoding to two observer's
    assessment. In green, the obtained VFoA using our aproach and in cyan, the
    mathing between them. Finally, the correspondent numbers with figure \ref{fig:bitmap}.}

>>>>

\usepackage{algorithm}
\usepackage{algpseudocode}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\columnwidth]{experimental_setup}
    \caption{\small \textbf{Experimental setting}. The areas of interest (AOI)
    considered of potential focuses of attention are circled in red.}
    \label{drawSetup}
\end{figure}

@@

The concept of \emph{with-me-ness} has been introduced in the field of
\emph{Computer Supported Collaborative Learning} (CSCL) by Sharma~\etal
in~\cite{sharma2014me}. Sharma~\etal introduce this concept in an attempt to
answer a recurrent teacher's question: {\it ``how much are the students with
me?''}. They distinguish what they call \emph{perceptual with-me-ness} (the
student follows what the teacher refers to with deictic gestures) from
\emph{conceptual with-me-ness} (the student follows what the teacher refers to
verbally), and they show in an eye-tracking study that \emph{conceptual
with-me-ness} in particular correlates with better learning performance. This
also relates to the concept of gaze cross-recurrence that has been shown to
reflect the quality of the interaction~\cite{jermann2012effects} in
collaborative learning tasks.

Sharma~\etal simply define \emph{conceptual with-me-ness} as the normalized
percentage of time during which the student's gaze overlapped the areas of
teaching slides currently referred to by the teacher.
In order to apply it to human-robot interactions, we propose to extend this
concept, and to define \emph{conceptual with-me-ness} as the normalized
ratio of time that the human interactant focuses its attention on the
attentional target expected by the robot for the current task (or sub-task).

\begin{algorithm}[h!]
    \centering

    \begin{algorithmic}[1]
    \Procedure{Compute With-me-ness} {}
    \State $d_w, d_e \gets 0$
    \State $t \gets t_{start}$
    \Repeat
    \If{$task(t) \neq$ nil \textbf{and} \par
    \hskip\algorithmicindent$F(task(t)) \neq \varnothing$ \textbf{and} \par
    \hskip\algorithmicindent $f(t) \neq$ nil } \label{algline:skiplosttrack}
    \If{$f(t) \in F(task(t))$}
            \State $d_w \gets d_w + 1$
        \EndIf
        \State $d_e \gets d_e + 1$
    \EndIf
    \State $t \gets t + 1$
    \Until{$t = t_{end}$}

    \State $\mathcal{W}_{[start, end]} \gets \frac{d_w}{d_e}$
    \State \Return $\mathcal{W}_{[start, end]}$
    \EndProcedure

    \end{algorithmic}

    \caption{\textbf{Computation of \emph{with-me-ness}}. $d_w$ stands for the duration
        the human is \emph{with} the robot, $d_e$ stands for the duration where
        the human would be \emph{expected to be with} the robot, $task(t)$
        represents the task performed by the robot at time $t$ (possibly none),
        $F(task)$ represents the (possibly empty) set of expected attentional
        targets associated to task $task$, $f(t)$ represents the actual focus of
        attention measured at time $t$. $\mathcal{W}_{[start, end]}$ represents
        the level of \emph{with-me-ness} from $t_{start}$ to $t_{end}$.}
    \label{alg:with-me-ness}
\end{algorithm}

Algorithm~\ref{alg:with-me-ness} provides a formal way of computing the level of
with-me-ness $\mathcal{W}$ between two time points $[t_{start}, t_{end}]$. A
notable difference with the original definition by Sharma~\etal is that, at a
given time $t$, the task $task(t)$ performed by the robot may elicit more than
one attentional target; thus, at a given time, more than one location can be
regarded as possible \emph{expected} focuses of attention for the human. For
example, a robot which is writing, could typically elicit gazes to its hand as
well as to its head. A human looking at either of these locations would be
considered to be \emph{with} the robot in terms of interaction. Also notable, we
exclude from the computation of $\mathcal{W}$ all of the periods of time where
the user's focus of attention can not be estimated (Algorithm~\ref{alg:with-me-ness},
line~\ref{algline:skiplosttrack}, third condition), typically because the user's
face is not visible at those times.

\begin{figure*}
    \centering
    \includegraphics[width=1.81\columnwidth]{realCaptured}
    \caption{\small Real VFoA of subject 4 accoding to two observer's
    assessment. In green, the obtained VFoA using our approach and in cyan, the
    matching between them. Finally, the correspondent numbers with figure \ref{fig:bitmap}.}
